---
title: "Kothari_Assignment1"
author: "Ravi Kothari"
date: "September 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment R and SQL

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates,
imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results
(observations) and store them in a SQL database. Load the information into an R dataframe.

## Description

For this assignment I created a database msda_607 with three tales.

- Movie
- rating
- Person

You can create the Database and tables with the SQL script on this link

## Load Datasets
```{r load data, eval=FALSE}
getwd()
setwd("C:/TEMP/")
df <- read.csv("movies.csv", header = FALSE) 
is.data.frame(df)
head(df)
names(df)
str(df)
names(df) <- c("title", "release_year", "genre", "length", "rating", "person_name", "age", "gender", "relation")
attach(df)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
