---
title: "Kothari_project1"
author: "Ravi Kothari"
date: "September 25, 2016"
output: html_document
---
## Setup

The following R packages are required for this project:

1. Downloader
2. Stringr

```{r setup}
library(downloader)
library(stringr)
```

## Project Description

In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents For the first player, the information would be:

Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents' ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

The goal of this project is to generate a .CSV file (which could for example be imported into a SQL database) with the following information for all of the players contained in the chess tournament cross table:

* Player's Name
* Player's State
* Total Number of Points
* Player's Pre-Rating
* Average Pre Chess Rating of Opponents

## Importing the dataset from the text file

Download the file, called `tournamentinfo.txt', from the GitHub URL. This file will be downloaded to your local working directory.

https://raw.githubusercontent.com/ravi-kothari/DATA-607/master/tournamentinfo.txt?token=AQGrklbb_mgZp2MfY-1Ixw30fsRR2reTks5X8gcVwA%3D%3D

```{r import_data, eval=TRUE}

url <- "https://raw.githubusercontent.com/ravi-kothari/DATA-607/master/tournamentinfo.txt?token=AQGrklbb_mgZp2MfY-1Ixw30fsRR2reTks5X8gcVwA%3D%3D"

download_file <- "tournamentinfo.txt" 

# the file will be downloaded to the working directory
downloader::download(url, download_file)

getwd()
```

## Extraction of the columns using regular expressions

```{r extract_columns, eval=TRUE}
library(stringr)
results <- readLines("tournamentinfo.txt")

# extract player name from the text file using regular expression
player_name <- unlist(str_extract_all(results, "\\w+[^USCF|a-z] ?\\w+ \\w+"))

# extract player_state
player_state <- unlist(str_extract_all(results, "(?:^|\\W)ON | MI | OH(?:$|\\W)"))

# extract total number of points for the player.

player_pts <- unlist(str_extract_all(results, "\\d\\.\\d"))

# extract pre rating for the player
# I divided the steps into couple of steps for this.

pre_rating_step1 <-  unlist(str_extract_all(results, "(R:\\s*)(\\d+)"))
pre_rating_step1[1]
pre_rating <- unlist(str_extract_all(pre_rating_step1, "(\\d+)"))
is.vector(pre_rating)
pre_rating <- as.numeric(pre_rating) # converting strings to numeric
pre_rating[1]


# For avg rating I divided the extraction into three steps

# Step 1: extract the full line of text that starts with "|" and a number.
avg_calc_step1 <- unlist(str_extract_all(results, "\\|[0-9].*"))

avg_calc_step1[16]


# used str_replace to replace empty spaces with 00 in the data.
avg_calc_step2 <- str_replace_all(avg_calc_step1, "\\s{1,2}\\|","00|")

avg_calc_step2[16]

# Step three is to only extract the number of opponent for each round.
avg_calc_step3 <- (str_extract_all(avg_calc_step2, "\\s\\d{1,2}"))

avg_calc_step3[16]

# Calculation of the Average Pre Chess Rating of Opponents

#converting the list into a matrix 
matrix_calc <- matrix(unlist(avg_calc_step3), byrow=TRUE, nrow=length(avg_calc_step3) )
dim(matrix_calc)
new_matrix <- t(apply(matrix_calc, 1,as.numeric)) 
dim(new_matrix)

#replacing each element in the matrix with actual rating for opponents and adding NA if its 0
for (i in 1:nrow(new_matrix)) 
  {
  for (j in 1:ncol(new_matrix))
    {
    if (new_matrix[i,j] == 0){
      new_matrix[i,j] = NA
    }
    else{
      new_matrix[i,j] <- pre_rating[new_matrix[i,j]]
    }
  }
}
# Finally calculating the mean of each row to get the average opponent ratings.
avg_opponents <- c(rowMeans(new_matrix, na.rm = TRUE))

# Putting everything together!
df <- data.frame(player_name,player_state, player_pts,pre_rating,avg_opponents)
colnames(df) <- c("Players Name","Players State", "Total Points", "Players Pre-Rating", "Opponents Ratings")

# Exporting the result into a CSV file.
write.csv(df, "Chess.csv", row.names=FALSE)
```

##Output
```{r output, echo=TRUE}
df
```



