---
title: "Data607_HW3"
author: "Ravi Kothari"
date: "September 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3 Copy the introductory example. The vector name stores the extracted names. 

R> name [1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy" [4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r load_dataset}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

first_name <- unlist(str_extract_all(name, "[:punct:] [[:alpha:]]{2,}$|[[:alpha:]]{2,} "))
first_name <- unlist(str_extract_all(first_name,"\\w+"))
first_name

last_name <- unlist(str_extract_all(name, "[^[:punct:]] [[:alpha:]]{2,}$|[[:alpha:]]{2,}, "))
last_name <- unlist(str_extract_all(last_name,"[[:alpha:]]{2,}"))

data.frame(first_name, last_name)
```

(B) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r 3B}
str_detect(name, unlist(str_extract_all(name, "[[:alpha:]]{2,}\\.")))
```

## 4 Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

(a) [0-9]+\$

```{r a}
example1 <- c("abcdefh$", "12456$45", "ads$afsa", "1234$abcd34$")
unlist(str_extract_all(example1,"[0-9]+\\$"))

```
String of numbers[0-9] followed by "$" sign. 

(b) \\b[a-z]{1,4}\\b

```{r b}
example2 <- c("tiny", "adasgvaf123", "aafdafa$", "also", "two")
unlist(str_extract_all(example2,"\\b[a-z]{1,4}\\b"))

```

the query above extracts [a-z] off of edges of the words of length 1 but not more than 4 characters. So the words should be of length 4 or less but consecutive.

(c) .*?\.txt$

```{r c}
example3 <- c("tiny", "ravi..txt", "123.txt", "234txt", "345.txt")
unlist(str_extract_all(example3,".*?\\.txt$"))

```

The above query will return anything which ends with .txt 

(d) \\d{2}/\\d{2}/\\d{4}

```{r d}
example4 <- c("tiny", "123445", "133/45/12333", "044/333/1984", "34/")
unlist(str_extract_all(example4,"\\d{2}/\\d{2}/\\d{4}"))

```

The above query is to extract any two digit just before a "/" then again exactly 2 digits followed by a "/" and then first four digits after the  second "/"" 


(e) <(.+?)>.+?</\1>

```{r e}
example5 <- c("<html>Text</html>", "123445", "133/45/12333", "044/333/1984", "34/")
unlist(str_extract_all(example5,"<(.+?)>.+?</\\1>"))

```

The above query extracts the opening and closing tags like html tags. The opening and closing should match followed including a "/" for closing tags.


## 9 The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.


clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

```{r 9}
secret <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

message <- unlist(str_extract_all(secret, "[[:upper:][[:punct:]]]"))
(str_replace_all(paste(message, collapse = ''), "[.]", " "))
```
```